Benchmarking FDR Methods - Summarizing Simulation Results (Part II)

### 2017-12-10 datasets/simulations/analysis-03-plot-tsim-summaries.Rmd
Update general plots of tsim/psim results

## parse each setting-block separately and save metrics in single file
for (dd in dl) {
    if (!file.exists(file.path(dd, "results-performanceMetrics-alpha05.rds"))) {
        fl <- list.files(dd, pattern=".rds", full.names=TRUE)
        tsbl <- list()
        for (f in fl) {
            ## read process data
            res <- readRDS(f)
            tsb <- plotsim_standardize(res, alpha = 0.05)
            tsbs <- tsb %>%
                filter(!(blabel %in% c("bl-df02", "bl-df04", "bl-df05", 
                                       paste0("ihw-a0", c(1:4, 6:9)), "ihw-a10")),
                       alpha == 0.05) %>%
                select(blabel, performanceMetric, value, rep)
            tsbl[[gsub(".*/(.*)\\.rds", "\\1", f)]] <- tsbs
        }
        tsbc <- as.tibble(bind_rows(tsbl, .id = "param"))
        saveRDS(tsbc, file.path(dd, "results-performanceMetrics-alpha05.rds"))
    }
}
```

@@

For simulation settings where the effect size (expected group differece) is not varying, a fixed difference of `1.5` was used for all alternative hypotheses. For simulation settings with the uniform (un-informative) or SE informative covariate, if the proportion of null hypotheses is not varying, `80%` of hypotheses were simulated under the null. For simulation settings with the "Boca-Leek" style informative covariate, a functional relationship was first defined between the probability of a test being null and the "informative covariate". Then, the informative covariates were sampled from Uniform(0, 1) and used to randomly assign each test as null or alternative according to the predefined functional relationship. The form of the relationship used here was taken from Setting II of the Boca-Leek simulations.

@@

```{r}
re_blabel <- function(tab) {
    tsb %>%
        mutate(blabel = fct_recode(f = blabel, "IHW" = "ihw-a05", "ASH" = "ashs",
                                   "Boca-Leek" = "bl-df03", "local FDR" = "lfdr",
                                   "FDRreg (theoretical)" = "scott-theoretical",
                                   "FDRreg (empirical)" = "scott-empirical",
                                   "Benjamini-Hochberg" = "bh",
                                   "Bonferroni" = "bonf", "q-value" = "qvalue"),
               takes_cov = blabel %in%
                   c("IHW", "ASH", "Boca-Leek", "FDRreg (theoretical)", "FDRreg (empirical)")) %>%
        mutate(blabel = factor(blabel, levels = lab_order))
}
```

>>>>

## parse each setting-block separately and save metrics in single file
for (dd in dl) {
    if (!file.exists(file.path(dd, "results-performanceMetrics-alpha05.rds"))) {
        print(paste0("running: ", basename(dd))) 
        fl <- list.files(dd, pattern=".rds", full.names=TRUE)
        tsbl <- list()
        for (f in fl) {
            ## read process data
            res <- readRDS(f)
            tsb <- plotsim_standardize(res, alpha = 0.05)
            tsbs <- tsb %>%
                filter(!(blabel %in% c("bl-df02", "bl-df04", "bl-df05", 
                                       paste0("ihw-a0", c(1:4, 6:9)), "ihw-a10")),
                       alpha == 0.05) %>%
                select(blabel, performanceMetric, value, rep)
            tsbl[[gsub(".*/(.*)\\.rds", "\\1", f)]] <- tsbs
        }
        tsbc <- as.tibble(bind_rows(tsbl, .id = "param"))
        print(paste0("saving: ", basename(dd))) 
        saveRDS(tsbc, file.path(dd, "results-performanceMetrics-alpha05.rds"))
        print(paste0("done: ", basename(dd))) 
    } else {
        print(paste0("skipping: ", basename(dd)))
    }
}
```

@@

For simulation settings where the effect size (expected group differece) is not varying, a fixed difference of `1.5` was used for all alternative hypotheses. For simulation settings with the uniform (uninformative) or SE (informative) covariate, if the proportion of null hypotheses is not varying, `80%` of hypotheses were simulated under the null. For simulation settings with the "Boca-Leek" style informative covariate, a functional relationship was first defined between the probability of a test being null and the "informative covariate". Then, the informative covariates were sampled from Uniform(0, 1) and used to randomly assign each test as null or alternative according to the predefined functional relationship. The form of the relationship used here was taken from Setting II of the Boca-Leek simulations. The marginal proportion of null hypothses is approximately `92.6%`. For the Manual Covariates, as with the Boca-Leek style covariate, a functional relationship was defined between the randomly sampled covariate and the probability of a hypothesis being null or alternative. Three functional relationships were tested: "step (less)", "step (more)", and "cubic". The first two specified a piecewise constant relationship of varying informativeness (more or less), and the third was a (continuous) scaled and shifted cubic function. The three variations on the manual covariate were chosen to have a marginal proportion of null hypotheses of `80%`. By definition, the ASH method makes use of the standard error (SE) for each hypothesis test, and does not accept any other additional information. The results for the ASH method are therefore **always** based on using the SE as the "informative covariate."

@@

```{r}
re_blabel <- function(tab) {
    tsb %>%
        mutate(blabel = fct_recode(f = blabel, "IHW" = "ihw-a05", "ASH" = "ashs",
                                   "Boca-Leek" = "bl-df03", "local FDR" = "lfdr",
                                   "FDRreg (theoretical)" = "scott-theoretical",
                                   "FDRreg (empirical)" = "scott-empirical",
                                   "Benjamini-Hochberg" = "bh",
                                   "Bonferroni" = "bonf", "q-value" = "qvalue"),
               takes_cov = blabel %in%
                   c("IHW", "ASH", "Boca-Leek", "local FDR", 
                     "FDRreg (theoretical)", "FDRreg (empirical)")) %>%
        mutate(blabel = factor(blabel, levels = lab_order))
}
```

### 2017-12-10 datasets/simulations/analysis-04-plot-psim-summaries.Rmd
Update general plots of tsim/psim results

This analysis is similar to the one presented in `analysis-03-plot-tsim-summaries.[Rmd|html]`. However, rather than looking at the results of two-sample t-tests, here, we consider simulations performed directly at the level of effect sizes and p-values without any underlying data. This approach to simulating data follows similar simulations carried out in the Boca-Leek, IHW ("size investing strategy" simulations), and ASH papers. Different simulation settings were considered to evaluate the impact of varying effect sizes, effect size distributions, effect size sampling distributions, proportions of null hypotheses, and presence or absence of an informative covariate. More details on the simulation settings considered are included in the header of the `psim-core-settings.R` script. Each setting was replicated 100 times.

Since these simulations do not rely on underlying simulated data, but rather simulate summary statistics directly, there is no way of calculating a corresponding standard error (SE) for each hypothesis test. *As a result, the ASH method, which requires the SE for each test, is not included in these simulations.*

@@

```{r}
lab_order <- c("Bonferroni", "Benjamini-Hochberg", "q-value", "local FDR",
               "IHW", "Boca-Leek", "FDRreg (theoretical)", "FDRreg (empirical)",
               "unadjusted")
```

@@

```{r}
re_blabel <- function(tab) {
    tsb %>%
        mutate(blabel = fct_recode(f = blabel, "IHW" = "ihw-a05",
                                   "Boca-Leek" = "bl-df03", "local FDR" = "lfdr",
                                   "FDRreg (theoretical)" = "scott-theoretical",
                                   "FDRreg (empirical)" = "scott-empirical",
                                   "Benjamini-Hochberg" = "bh",
                                   "Bonferroni" = "bonf", "q-value" = "qvalue"),
               takes_cov = blabel %in%
                   c("IHW", "Boca-Leek", "FDRreg (theoretical)", "FDRreg (empirical)")) %>%
        mutate(blabel = factor(blabel, levels = lab_order))
}
```

@@

@@

```{r}
## older run with alternative tests at 1.5.
f <- "data-psim/M100/_pi0-uniform-delta15/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

@@

@@

@@

>>>>

This analysis is similar to the one presented in `analysis-03-plot-tsim-summaries.[Rmd|html]`. However, rather than looking at the results of two-sample t-tests, here, we consider simulations performed directly at the level of effect sizes and p-values without any underlying data. This approach to simulating data follows similar simulations carried out in the Boca-Leek, IHW ("size investing strategy" simulations), and ASH papers. Different simulation settings were considered to evaluate the impact of varying effect sizes, effect size distributions, sampling distributions, proportions of null hypotheses, and presence or absence of an informative covariate. By definition, the ASH method makes use of the standard error (SE) for each hypothesis test, and does not accept any other additional information. The results for the ASH method are therefore always based on using the SE as the "informative covariate." More details on the simulation settings considered are included in the header of the `psim-core-settings.R` script. Each setting was replicated 100 times.

Since these simulations do not rely on underlying simulated data, but rather simulate summary statistics directly, there is no way of calculating a corresponding SE for each hypothesis test. *As such, the ASH method, which requires the SE for each test, is run using the true (typically unknown) SE underlying the simulation.* Therefore, the ASH results must be interpreted with caution.

@@

```{r}
lab_order <- c("Bonferroni", "Benjamini-Hochberg", "q-value", "local FDR",
               "IHW", "Boca-Leek", "FDRreg (theoretical)", "FDRreg (empirical)",
               "ASH", "unadjusted")
```

@@

```{r}
re_blabel <- function(tab) {
    tsb %>%
        mutate(blabel = fct_recode(f = blabel, "IHW" = "ihw-a05", "ASH" = "ashs",
                                   "Boca-Leek" = "bl-df03", "local FDR" = "lfdr",
                                   "FDRreg (theoretical)" = "scott-theoretical",
                                   "FDRreg (empirical)" = "scott-empirical",
                                   "Benjamini-Hochberg" = "bh",
                                   "Bonferroni" = "bonf", "q-value" = "qvalue"),
               takes_cov = blabel %in%
                   c("IHW", "ASH", "Boca-Leek", "local FDR",
                     "FDRreg (theoretical)", "FDRreg (empirical)")) %>%
        mutate(blabel = factor(blabel, levels = lab_order))
}
```

@@

<!-- ######################################################################## -->
### Manual Covariate (Step - Less Informative)

```{r}
f <- "data-psim/M100/esize_fixed-bl-step-less/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## summarize results in plot
gen_plot("Direct Parameter Simulation w/ Varying Effect Size, Manual Covariate (step - less informative) (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### Manual Covariate (Step - More Informative)

```{r}
f <- "data-psim/M100/esize_fixed-bl-step-more/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## summarize results in plot
gen_plot("Direct Parameter Simulation w/ Varying Effect Size, Manual Covariate (step - more informative) (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### Manual Covariate (Cubic)

```{r}
f <- "data-psim/M100/esize_fixed-bl-cubic/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## summarize results in plot
gen_plot("Direct Parameter Simulation w/ Varying Effect Size, Manual Covariate (cubic) (100 replications)",
         tsb, cutoffs)
```

@@

```{r}
f <- "data-psim/M100/pi0-uniform/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

@@

<!-- ######################################################################## -->
### Manual Covariate (Step - Less Informative)

```{r}
f <- "data-psim/M100/esize_random_ua-bl-step-less/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## summarize results in plot
gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled UA Effect, Manual Covariate (step - less informative) (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### Manual Covariate (Step - More Informative)

```{r}
f <- "data-psim/M100/esize_random_ua-bl-step-more/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## summarize results in plot
gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled UA Effect, Manual Covariate (step - more informative) (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### Manual Covariate (Cubic)

```{r}
f <- "data-psim/M100/esize_random_ua-bl-cubic/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## summarize results in plot
gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled UA Effect, Manual Covariate (cubic) (100 replications)",
         tsb, cutoffs)
```

@@

<!-- ######################################################################## -->
### Manual Covariate (Step - Less Informative)

```{r}
f <- "data-psim/M100/esize_random_shift-bl-step-less/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## summarize results in plot
gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled Mean Shifted Effect, Manual Covariate (step - less informative) (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### Manual Covariate (Step - More Informative)

```{r}
f <- "data-psim/M100/esize_random_shift-bl-step-more/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## summarize results in plot
gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled Mean Shifted Effect, Manual Covariate (step - more informative) (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### Manual Covariate (Cubic)

```{r}
f <- "data-psim/M100/esize_random_shift-bl-cubic/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## summarize results in plot
gen_plot("Direct Parameter Simulation w/ Varying Randomly Sampled Mean Shifted Effect, Manual Covariate (cubic) (100 replications)",
         tsb, cutoffs)
```

@@

<!-- ######################################################################## -->
### Manual Covariate (Step - Less Informative)

```{r}
f <- "data-psim/M100/altnoise-bl-step-less/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## summarize results in plot
gen_plot("Direct Parameter Simulation w/ Varying Sampling Distribution, Manual Covariate (step - less informative) (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### Manual Covariate (Step - More Informative)

```{r}
f <- "data-psim/M100/altnoise-bl-step-more/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## summarize results in plot
gen_plot("Direct Parameter Simulation w/ Varying Sampling Distribution, Manual Covariate (step - more informative) (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
### Manual Covariate (Cubic)

```{r}
f <- "data-psim/M100/altnoise-bl-cubic/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## summarize results in plot
gen_plot("Direct Parameter Simulation w/ Varying Sampling Distribution, Manual Covariate (cubic) (100 replications)",
         tsb, cutoffs)
```

<!-- ######################################################################## -->
## Varying Parameter Sampling Distribution (all null)
<!-- ######################################################################## -->

Similar simulation settings as above, where various sampling distributions are considered, but with `pi0 = 1`, i.e. all null hypotheses.

```{r}
f <- "data-psim/M100/allnull-uniform/results-performanceMetrics-alpha05.rds"
tsb <- readRDS(f)

## expected thresholds based on simulation setting
cutoffs <- data.frame(performanceMetric = factor(c("FPR", "rejectprop"),
                                                 levels=levels(tsb$performanceMetric)),
                      expected = c(0.05, 0.00))

## summarize results in plot
gen_plot("Direct Parameter Simulation w/ Varying Sampling Dist (all null), Non-Informative Covariate (100 replications)",
         tsb, cutoffs)
```
